### 2013-09-28 complexity.tex
first docs

\section{Complexity}
\label{section:complexity}
\textit{Q5. Complexity: What is the complexity of an artifact under review and what is the simplest way of measuring this complexity?}\\

Contributions are sent to the mailing list in the form of diffs for review.
Measuring the complexity of a change differs from measuring the complexity of a
file or the entire system \cite{German2005SM}. For example, changing a large
number of simple functions across multiple files may be as difficult as
changing the code contained within a single complex function.
%
Traditional complexity measures often require the whole file or system and were
not designed to measure the complexity of code fragments \cite{Hindle2008ICPC}.
Further, multiple studies have found that these measures correlate very
strongly with the lines of code in a file
\cite{Graves2000IEEE,Leszak2002JSS,Herraiz2007MSR} and may not provide much
additional information about the complexity of the system.

In this section, we provide various possible measures of change complexity.
Our measures are applied to patch contributions~\cite{German2005SM}.
%
All of our measures are based upon the premise that changes that are farther
from each other or involve multiple entities are more complex than those closer
together involving fewer entities. 
%
In total, we measure change complexity in seven ways:

\begin{itemize}

\item  the churn or change size (see previous section \ref{section:size}),

\item the number of modified files in a diff,

\item the number of diffs per review,

\item the number of and distance between contiguous change blocks within a diff
(\ie hunks),

\item the directory distance between these files, and

\item depth of indentation of a change.

\end{itemize}

%
Since the goal of each measure is the same, to assess the complexity of a
change, we expect some of our measures to be highly correlated. 
%we want the simpliest most parsimonious measures. 
In the interest of parsimony, we will select the simplest set of measures that
adequately captures change complexity. Future work could examine each of these
measure outside of the context of peer review. Given the goals of this work,
when appropriate, instead of providing a box plot, we show that a more complex
measure correlates with a simpler measure\footnote{We use non-parametric
Spearman correlations, correlation are statistically significant at $ p < .001$
unless otherwise specified}. In this case, the complex measure is dropped from
further analysis.

{\bf Files and diffs:} we count the number of files contained in a review and
the number of diffs. The posting of a diff often represents a re-write of the
original contribution or a sequence of diffs that implement a larger feature.
As more files change and more diffs are produced, reviewers will likely need an
understanding of a larger proportion of the system.
%-- see discussion below and Figures \ref{fig:rtc_files} and
%\ref{fig:ctr_files}.

Figures \ref{fig:rtc_files} and \ref{fig:ctr_files} show the number of files
modified per review. For {\bf RTC}, there are one to two files modified in the
median case and between three and five at the $75^{th}$ percentile.  For {\bf
CTR}, there is one and between three and four for the median and $75^{th}$
percentile, respectively. There is a moderate correlation between churn and the
number of files changed (min (kde ctr) $= .56$, mean $= .64$, max (lx)
$=.74$). Since Linux is the only project that shows a strong correlation with
churn, we keep the number of changed files as a predictor, but also add an
interaction term between churn and number of files (See models in Section
\ref{section:model}).

\begin{figure}
\centering
\includegraphics[width=\plotImageWidth]{figures/rtc/files} 
\caption{RTC -- Number of modified files per contribution}
\label{fig:rtc_files}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=\plotImageWidth]{figures/ctr/files}
\caption{CTR -- Number of modified files per contribution}
\label{fig:ctr_files}
\end{figure}

The number of diffs per review is low. For {\bf CTR} at the $90^{th}$
percentile it is one, indicating that few subsequent diffs are posted during
CTR. For {\bf RTC} it is one at the median and between one and two at the
$75^{th}$ percentile, indicating that posting updated diffs is also rare with
RTC. The correlation between the number of diffs and the number of files is
weak (min (kde ctr) $=.11$, mean $= .26$, max (lx) $= .57$) as is the
correlation with churn (min (kde ctr) $= .10$, mean $= .25$, max (lx) $= .54$).
The number of diffs is kept as a predictor.

{\bf Change blocks:} we measure the number of contiguous change blocks, or
hunks, and the distance between these blocks in a modified file and sum them
across the entire review.  Contiguous changes  are likely easier to review than
changes that are far apart and that are potentially in multiple functions
within a file.  
%
We find that the number of change blocks correlates strongly with a simpler
measure, churn (min (kde) $= .70$, mean $= .81$, max (gn) $= .86$).
%Just c,cpp (mean $= .79$, kde $= .69$ to svn $= .84$).
Intuitively and in practice, it appears that as the number of changed blocks of
code increases, so to does the number of total changes, or churn. The distance
between change blocks is dropped from further analysis, but could be
investigated in future work with different objectives.

%
{\bf Directory distance:} we measure the directory distance between the files
in a review. The premise for this measure is that files that perform similar
functions are typically closer in the directory hierarchy than files that
perform dissimilar functions~\cite{Bowman1999ICSE}, thus the directory
structure loosely mirrors the system architecture~\cite{RigbySubTOSEM}. The
distance of two files in the same directory have a distance of zero, while the
distance for files in different directories is the number of directories
between the two files in the directory hierarchy.  We expect that changes that
involve files that are far apart will crosscut the functionality of a system
and be more complex to review.
%
We find that the directory distance of the files in a change correlates
strongly with a simpler measure, the total number of modified files in a review
(min (fb rtc $= .78$, mean $= .87$ max (fb ctr) $= .92$).  Intuitively we can
see that as the number of files increases, the chance that two files will be
further apart in the directory structure also increases.

{\bf Indentation:} Hindle \etal \cite{Hindle2008ICPC} created a complexity
measure that can be used to measure the complexity of the entire system or of a
change by examining the level of indentation on source lines. We calculate
their measure for reviews. They found that the strongest predictors of
complexity were the sum and standard deviation of the indentation of source
lines. Since our contribution sizes are so small, see Section
\ref{section:size}, the standard deviation is not meaningful, so we just sum
the amount of indentation.
%
%Furthermore, their measure is dependent on the norms in the project and the
%programming language.  as different levels of indentation may be used.  indent
The sum of indentation correlates strongly with churn (min (fb) $= .78$, mean
$= .89$, max (lx) $= .94$), so it is dropped from further analysis.

When one or more measures are highly correlated, parsimony requires us to drop
the more complex measures in our model. Churn, see Section \ref{section:size},
is kept over indentation and number of change blocks, and the number of
modified files in a change is kept over directory distance of files in a
change.

In conclusion, it is intuitive that a large amount of churn and modified files
would affect a large number and diverse sections of the system and thus
represent a complex change. It is also reasonable that the more diffs generated
during a review the more complex the change.
%
In agreement with intuition and with previous literature comparing other
complexity measures with the number of lines in a file or system, we find that
the number of modified files and the churn correlate strongly with other
proposed measures of complexity and seem to provide a reasonable proxy for the
complexity of a review. Thus, we keep these two simple measures of complexity
and the number of diffs per review, and we discard the other four more complex,
\ie more difficult to calculate, measures.

### 2013-10-07 paper/old_stuff/rejected_text.tex
Intro, research questions, and outline

--- 2013-09-28 complexity.tex
@@
-\section{Complexity}
-\label{section:complexity}
-\textit{Q5. Complexity: What is the complexity of an artifact under review and what is the simplest way of measuring this complexity?}\\
+The complexity of a file or a system is not the same as the complexity of the
+changes made to the system. For example, a change to the error message in an
+otherwise complex file, represents a simple change in a complex file.
 
-Contributions are sent to the mailing list in the form of diffs for review.
-Measuring the complexity of a change differs from measuring the complexity of a
-file or the entire system \cite{German2005SM}. For example, changing a large
-number of simple functions across multiple files may be as difficult as
-changing the code contained within a single complex function.
-%
+The goal of this work, is to develop measures of the complexity of a change.
+
+A change to the eA change that modifies the output message shown to users
+
 Traditional complexity measures often require the whole file or system and were
@@
 
-In this section, we provide various possible measures of change complexity.
-Our measures are applied to patch contributions~\cite{German2005SM}.
-%
-All of our measures are based upon the premise that changes that are farther
-from each other or involve multiple entities are more complex than those closer
-together involving fewer entities. 
-%
-In total, we measure change complexity in seven ways:
-
-\begin{itemize}
-
-\item  the churn or change size (see previous section \ref{section:size}),
-
-\item the number of modified files in a diff,
-
-\item the number of diffs per review,
-
-\item the number of and distance between contiguous change blocks within a diff
-(\ie hunks),
-
-\item the directory distance between these files, and
-
-\item depth of indentation of a change.
-
-\end{itemize}
-
-%
-Since the goal of each measure is the same, to assess the complexity of a
-change, we expect some of our measures to be highly correlated. 
-%we want the simpliest most parsimonious measures. 
-In the interest of parsimony, we will select the simplest set of measures that
-adequately captures change complexity. Future work could examine each of these
-measure outside of the context of peer review. Given the goals of this work,
-when appropriate, instead of providing a box plot, we show that a more complex
-measure correlates with a simpler measure\footnote{We use non-parametric
-Spearman correlations, correlation are statistically significant at $ p < .001$
-unless otherwise specified}. In this case, the complex measure is dropped from
-further analysis.
-
-{\bf Files and diffs:} we count the number of files contained in a review and
-the number of diffs. The posting of a diff often represents a re-write of the
-original contribution or a sequence of diffs that implement a larger feature.
-As more files change and more diffs are produced, reviewers will likely need an
-understanding of a larger proportion of the system.
-%-- see discussion below and Figures \ref{fig:rtc_files} and
-%\ref{fig:ctr_files}.
-
-Figures \ref{fig:rtc_files} and \ref{fig:ctr_files} show the number of files
-modified per review. For {\bf RTC}, there are one to two files modified in the
-median case and between three and five at the $75^{th}$ percentile.  For {\bf
-CTR}, there is one and between three and four for the median and $75^{th}$
-percentile, respectively. There is a moderate correlation between churn and the
-number of files changed (min (kde ctr) $= .56$, mean $= .64$, max (lx)
-$=.74$). Since Linux is the only project that shows a strong correlation with
-churn, we keep the number of changed files as a predictor, but also add an
-interaction term between churn and number of files (See models in Section
-\ref{section:model}).
-
-\begin{figure}
-\centering
-\includegraphics[width=\plotImageWidth]{figures/rtc/files} 
-\caption{RTC -- Number of modified files per contribution}
-\label{fig:rtc_files}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=\plotImageWidth]{figures/ctr/files}
-\caption{CTR -- Number of modified files per contribution}
-\label{fig:ctr_files}
-\end{figure}
-
-The number of diffs per review is low. For {\bf CTR} at the $90^{th}$
-percentile it is one, indicating that few subsequent diffs are posted during
-CTR. For {\bf RTC} it is one at the median and between one and two at the
-$75^{th}$ percentile, indicating that posting updated diffs is also rare with
-RTC. The correlation between the number of diffs and the number of files is
-weak (min (kde ctr) $=.11$, mean $= .26$, max (lx) $= .57$) as is the
-correlation with churn (min (kde ctr) $= .10$, mean $= .25$, max (lx) $= .54$).
-The number of diffs is kept as a predictor.
-
-{\bf Change blocks:} we measure the number of contiguous change blocks, or
-hunks, and the distance between these blocks in a modified file and sum them
-across the entire review.  Contiguous changes  are likely easier to review than
-changes that are far apart and that are potentially in multiple functions
-within a file.  
-%
-We find that the number of change blocks correlates strongly with a simpler
-measure, churn (min (kde) $= .70$, mean $= .81$, max (gn) $= .86$).
-%Just c,cpp (mean $= .79$, kde $= .69$ to svn $= .84$).
-Intuitively and in practice, it appears that as the number of changed blocks of
-code increases, so to does the number of total changes, or churn. The distance
-between change blocks is dropped from further analysis, but could be
-investigated in future work with different objectives.
-
-%
-{\bf Directory distance:} we measure the directory distance between the files
-in a review. The premise for this measure is that files that perform similar
-functions are typically closer in the directory hierarchy than files that
-perform dissimilar functions~\cite{Bowman1999ICSE}, thus the directory
-structure loosely mirrors the system architecture~\cite{RigbySubTOSEM}. The
-distance of two files in the same directory have a distance of zero, while the
-distance for files in different directories is the number of directories
-between the two files in the directory hierarchy.  We expect that changes that
-involve files that are far apart will crosscut the functionality of a system
-and be more complex to review.
-%
-We find that the directory distance of the files in a change correlates
-strongly with a simpler measure, the total number of modified files in a review
-(min (fb rtc $= .78$, mean $= .87$ max (fb ctr) $= .92$).  Intuitively we can
-see that as the number of files increases, the chance that two files will be
-further apart in the directory structure also increases.
-
-{\bf Indentation:} Hindle \etal \cite{Hindle2008ICPC} created a complexity
-measure that can be used to measure the complexity of the entire system or of a
-change by examining the level of indentation on source lines. We calculate
-their measure for reviews. They found that the strongest predictors of
-complexity were the sum and standard deviation of the indentation of source
-lines. Since our contribution sizes are so small, see Section
-\ref{section:size}, the standard deviation is not meaningful, so we just sum
-the amount of indentation.
-%
-%Furthermore, their measure is dependent on the norms in the project and the
-%programming language.  as different levels of indentation may be used.  indent
-The sum of indentation correlates strongly with churn (min (fb) $= .78$, mean
-$= .89$, max (lx) $= .94$), so it is dropped from further analysis.
-
-When one or more measures are highly correlated, parsimony requires us to drop
-the more complex measures in our model. Churn, see Section \ref{section:size},
-is kept over indentation and number of change blocks, and the number of
-modified files in a change is kept over directory distance of files in a
-change.
-
-In conclusion, it is intuitive that a large amount of churn and modified files
-would affect a large number and diverse sections of the system and thus
-represent a complex change. It is also reasonable that the more diffs generated
-during a review the more complex the change.
-%
-In agreement with intuition and with previous literature comparing other
-complexity measures with the number of lines in a file or system, we find that
-the number of modified files and the churn correlate strongly with other
-proposed measures of complexity and seem to provide a reasonable proxy for the
-complexity of a review. Thus, we keep these two simple measures of complexity
-and the number of diffs per review, and we discard the other four more complex,
-\ie more difficult to calculate, measures.
+System 
+Simple changes can be made to complex files and complex changes can be made to simple files. For example, a one line ch
+One
+point to note is that

### 2013-10-07 paper/paper.tex
Intro, research questions, and outline

--- 2013-09-28 complexity.tex
@@
-\section{Complexity}
-\label{section:complexity}
-\textit{Q5. Complexity: What is the complexity of an artifact under review and what is the simplest way of measuring this complexity?}\\
+\section{Introduction}
 
-Contributions are sent to the mailing list in the form of diffs for review.
-Measuring the complexity of a change differs from measuring the complexity of a
-file or the entire system \cite{German2005SM}. For example, changing a large
-number of simple functions across multiple files may be as difficult as
-changing the code contained within a single complex function.
-%
-Traditional complexity measures often require the whole file or system and were
-not designed to measure the complexity of code fragments \cite{Hindle2008ICPC}.
-Further, multiple studies have found that these measures correlate very
-strongly with the lines of code in a file
-\cite{Graves2000IEEE,Leszak2002JSS,Herraiz2007MSR} and may not provide much
-additional information about the complexity of the system.
+The complexity of a change can be very different from the complexity of the
+files or system that contain a change (e.g., a change to many simple files
+might be as complex as a change to one difficult file). Change complexity is
+related to, ``how hard is it to understand and review a change" vs system
+complexity, which measures for example, "how hard is it to understand and
+potentially modify the system".
 
-In this section, we provide various possible measures of change complexity.
-Our measures are applied to patch contributions~\cite{German2005SM}.
+Traditional complexity measures (\eg McCabe) are normed on or require a whole
+file or system and were not designed to measure the complexity of changes and
+code fragments. The goal of this paper is to evaluate which measures are the
+best indicators of the complexity the changes to an evolving system.
+
+\textbf{We have the following research questions:}
+
+\begin{enumerate}
+
+\item Does the difference of traditional complexity measures before and after a change correlate with change complexity measures?
+
+\item Does the weighted sum of change complexity measures up to a version correlated with the traditional complexity measures at that version?
+
+\end{enumerate}
+
+\section{Background and Measures}
+
+\subsection{Traditional Measures}
+TODO Describe in a few paragraphs each of the following metrics ...
+
+McCabe etc
+
+\subsection{Change Complexity Measures}
+
+Each change is calcuated on the difference between two versions of a file.
 %
@@
 
-\item  the churn or change size (see previous section \ref{section:size}),
+\item  the churn or change size,
 
-\item the number of modified files in a diff,
+\item the number of modified files per commit,
 
-\item the number of diffs per review,
+\item the number of diffs per commit,
 
@@
 
-\item the directory distance between these files, and
+\item the directory distance between files, and
 
@@
 In the interest of parsimony, we will select the simplest set of measures that
-adequately captures change complexity. Future work could examine each of these
-measure outside of the context of peer review. Given the goals of this work,
-when appropriate, instead of providing a box plot, we show that a more complex
-measure correlates with a simpler measure\footnote{We use non-parametric
-Spearman correlations, correlation are statistically significant at $ p < .001$
-unless otherwise specified}. In this case, the complex measure is dropped from
-further analysis.
+adequately captures change complexity.
 
-{\bf Files and diffs:} we count the number of files contained in a review and
-the number of diffs. The posting of a diff often represents a re-write of the
-original contribution or a sequence of diffs that implement a larger feature.
-As more files change and more diffs are produced, reviewers will likely need an
-understanding of a larger proportion of the system.
-%-- see discussion below and Figures \ref{fig:rtc_files} and
-%\ref{fig:ctr_files}.
-
-Figures \ref{fig:rtc_files} and \ref{fig:ctr_files} show the number of files
-modified per review. For {\bf RTC}, there are one to two files modified in the
-median case and between three and five at the $75^{th}$ percentile.  For {\bf
-CTR}, there is one and between three and four for the median and $75^{th}$
-percentile, respectively. There is a moderate correlation between churn and the
-number of files changed (min (kde ctr) $= .56$, mean $= .64$, max (lx)
-$=.74$). Since Linux is the only project that shows a strong correlation with
-churn, we keep the number of changed files as a predictor, but also add an
-interaction term between churn and number of files (See models in Section
-\ref{section:model}).
-
-\begin{figure}
-\centering
-\includegraphics[width=\plotImageWidth]{figures/rtc/files} 
-\caption{RTC -- Number of modified files per contribution}
-\label{fig:rtc_files}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=\plotImageWidth]{figures/ctr/files}
-\caption{CTR -- Number of modified files per contribution}
-\label{fig:ctr_files}
-\end{figure}
-
-The number of diffs per review is low. For {\bf CTR} at the $90^{th}$
-percentile it is one, indicating that few subsequent diffs are posted during
-CTR. For {\bf RTC} it is one at the median and between one and two at the
-$75^{th}$ percentile, indicating that posting updated diffs is also rare with
-RTC. The correlation between the number of diffs and the number of files is
-weak (min (kde ctr) $=.11$, mean $= .26$, max (lx) $= .57$) as is the
-correlation with churn (min (kde ctr) $= .10$, mean $= .25$, max (lx) $= .54$).
-The number of diffs is kept as a predictor.
+{\bf Files and diffs:} we count the number of files contained in a commit.
+The more files that change, the larger the affect proportion of the system.
+Commits with a large number of files will likely be difficult to understand.
 
@@
 within a file.
-%
-We find that the number of change blocks correlates strongly with a simpler
-measure, churn (min (kde) $= .70$, mean $= .81$, max (gn) $= .86$).
-%Just c,cpp (mean $= .79$, kde $= .69$ to svn $= .84$).
-Intuitively and in practice, it appears that as the number of changed blocks of
-code increases, so to does the number of total changes, or churn. The distance
-between change blocks is dropped from further analysis, but could be
-investigated in future work with different objectives.
 
-%
 {\bf Directory distance:} we measure the directory distance between the files
-in a review. The premise for this measure is that files that perform similar
+in a commit. The premise for this measure is that files that perform similar
 functions are typically closer in the directory hierarchy than files that
@@
 and be more complex to review.
-%
-We find that the directory distance of the files in a change correlates
-strongly with a simpler measure, the total number of modified files in a review
-(min (fb rtc $= .78$, mean $= .87$ max (fb ctr) $= .92$).  Intuitively we can
-see that as the number of files increases, the chance that two files will be
-further apart in the directory structure also increases.
 
@@
 complexity were the sum and standard deviation of the indentation of source
-lines. Since our contribution sizes are so small, see Section
-\ref{section:size}, the standard deviation is not meaningful, so we just sum
-the amount of indentation.
-%
-%Furthermore, their measure is dependent on the norms in the project and the
-%programming language.  as different levels of indentation may be used.  indent
-The sum of indentation correlates strongly with churn (min (fb) $= .78$, mean
-$= .89$, max (lx) $= .94$), so it is dropped from further analysis.
+lines.
 
-When one or more measures are highly correlated, parsimony requires us to drop
-the more complex measures in our model. Churn, see Section \ref{section:size},
-is kept over indentation and number of change blocks, and the number of
-modified files in a change is kept over directory distance of files in a
-change.
+%TODO weighted number of people who changed a file?
 
-In conclusion, it is intuitive that a large amount of churn and modified files
-would affect a large number and diverse sections of the system and thus
-represent a complex change. It is also reasonable that the more diffs generated
-during a review the more complex the change.
-%
-In agreement with intuition and with previous literature comparing other
-complexity measures with the number of lines in a file or system, we find that
-the number of modified files and the churn correlate strongly with other
-proposed measures of complexity and seem to provide a reasonable proxy for the
-complexity of a review. Thus, we keep these two simple measures of complexity
-and the number of diffs per review, and we discard the other four more complex,
-\ie more difficult to calculate, measures.
+\section{Methodology and Data}
+
+TODO: this is where you put your figure and discription
+
+\section{Results}
+
+In this section, we compare how correlated each measure is to each other.
+
+\section{Threats to validity}
+
+\section{Conclusion}

### 2013-10-19 paper/paper.tex
Added preamble and content to paper.tex

@@
+\documentclass[11pt,twoside,a4paper]{article}
+\usepackage{amssymb, amsmath}
+
+\begin{document}
 \section{Introduction}
@@
 
-McCabe etc
+\begin{enumerate}
+\item[Source Line of Code] To define SLOC we use the definition given by Conte [Conte 1986]:
+A line of code is any line of program text that is not a comment or blank line, regardless of the
+number of statements or fragments of statements on the line. This specifically includes all lines containing program headers, declarations, and executable and non-executable statements.
+
+\item[Lines of Code (LOC)]  Lines of code refers to the total number of lines in each source code file including comments and blank lines.
+It is a straightforward measure and can be calculated using many available tools.
+
+\item[McCabe's Cyclomatic Complexity]
+McCabe’s Cyclomatic complexity is one of the earliest complexity measures developed by Thomas J.
+McCabe in 1976.
+It directly measures the number of linearly independent paths through a program's source code.
+Any program can be represented as a graph with the simplest element being a flat series of statements with no conditions, loops or branches.
+[1] For a graph $G$ with $n$ vertices, $e$ edges and $p$ exit points, the complexity $v$ is defined as follows:
+
+\begin{equation}
+v(G)=e-n+2p
+\end{equation}
+
+It is worth mentioning that the minimum value for the cyclomatic complexity metric is 1, which corresponds to the flat series of statements with no bifurcations or loops.
+Every additional region in the flow graph would increase the Cyclomatic complexity by one unit.
+
+\end{enumerate}
 
@@
 \section{Conclusion}
+
+\end{document}

### 2013-10-22 paper/paper.tex
Added figure in .tex file, added macros for (e.g, i.e)

@@
 \documentclass[11pt,twoside,a4paper]{article}
-\usepackage{amssymb, amsmath}
+\usepackage{amssymb, amsmath,graphicx,soul,color}
+
+\newcommand*{\eg}{e.g. }
+\newcommand*{\ie}{i.e. }
 
@@
 The complexity of a change can be very different from the complexity of the
-files or system that contain a change (e.g., a change to many simple files
+files or system that contain a change (\eg, a change to many simple files
 might be as complex as a change to one difficult file). Change complexity is
@@
 \begin{enumerate}
-\item[Source Line of Code] To define SLOC we use the definition given by Conte [Conte 1986]:
+\item{\bf{Source Line of Code: }} To define SLOC we use the definition given by Conte [Conte 1986]:
 A line of code is any line of program text that is not a comment or blank line, regardless of the
@@
 
-\item[Lines of Code (LOC)]  Lines of code refers to the total number of lines in each source code file including comments and blank lines.
+\item{\bf{Lines of Code (LOC): }} Lines of code refers to the total number of lines in each source code file including comments and blank lines.
 It is a straightforward measure and can be calculated using many available tools.
 
-\item[McCabe's Cyclomatic Complexity]
+\item{\bf{McCabe's Cyclomatic Complexity: }}
 McCabe’s Cyclomatic complexity is one of the earliest complexity measures developed by Thomas J.
@@
 
-\item  the churn or change size,
+\item The churn or change size,
 
-\item the number of modified files per commit,
+\item The number of modified files per commit,
 
-\item the number of diffs per commit,
+\item The number of diffs per commit,
 
-\item the number of and distance between contiguous change blocks within a diff
+\item The number of and distance between contiguous change blocks within a diff
 (\ie hunks),
 
-\item the directory distance between files, and
+\item The directory distance between files, and
 
-\item depth of indentation of a change.
+\item The depth of indentation of a change.
 
@@
 %we want the simpliest most parsimonious measures. 
-In the interest of parsimony, we will select the simplest set of measures that
-adequately captures change complexity.
+In the interest of parsimony, we will select the simplest set of measures that adequately captures change complexity.
 
@@
 
-{\bf Indentation:} Hindle \etal \cite{Hindle2008ICPC} created a complexity
+{\bf Indentation:} Hindle \cite{Hindle2008ICPC} created a complexity
 measure that can be used to measure the complexity of the entire system or of a
@@
 
-TODO: this is where you put your figure and discription
+\hl{TODO: this is where you put your figure and discription}
+
+\begin{figure}[h!]
+  \centering
+   \includegraphics[width=1\textwidth]{change_complexity_figure}
+   \caption{"Change Complexity vs. Traditional Complexity"}
+\end{figure}
 

### 2013-11-18 paper/paper.tex
Changed width of figure
changed width of figure
added potential future work
Added description to figure

@@
-\documentclass[11pt,twoside,a4paper]{article}
+\documentclass[8pt,twocolumn,a4paper]{article}
 \usepackage{amssymb, amsmath,graphicx,soul,color}
@@
 
-\subsection{Traditional Measures}
-TODO Describe in a few paragraphs each of the following metrics ...
+\subsection{Traditional Complexity Measures}
 
@@
 
-\hl{TODO: this is where you put your figure and discription}
+In figure 1 we illustrate a hypothetical system which undergoes numerous changes.
+The vertical lines represent each version.
+During each version's development, the change complexity measures are assessed.
+In between versions, traditional complexity measures are computed.
+These are denoted by CCM and TCM respectively.
 
@@
   \centering
-   \includegraphics[width=1\textwidth]{change_complexity_figure}
+ %  \includegraphics[width=0.5\textwidth]{change_complexity_figure}
    \caption{"Change Complexity vs. Traditional Complexity"}
 \end{figure}
+
+\begin{equation}
+{{CCM_{v}(i)} \sim {TCM_{v(i+1)}}-{TCM_{v(i)}}}
+\end{equation}
+
+\begin{equation}
+{{TCM_{v(i+1)}} \sum\limits_{i=0}^iCCM_{v(i)}}
+\end{equation}
 
@@
 
+Future work: Does McCabe predict anything? Do the change complexity measures do a better job of prediction?
+
 \end{document}

### 2013-11-19 paper/paper.tex
Added remaining traditional complexity measures

@@
 \subsection{Traditional Complexity Measures}
+{\bf{\emph{Metrics Related to Lines of Code}}}
+\newline
+\begin{enumerate}
 
-\begin{enumerate}
-\item{\bf{Source Line of Code: }} To define SLOC we use the definition given by Conte [Conte 1986]:
+\item{\bf{Source Line of Code(SLOC): }} To define SLOC we use the definition given by Conte [Conte 1986]:
 A line of code is any line of program text that is not a comment or blank line, regardless of the
@@
 
-\item{\bf{McCabe's Cyclomatic Complexity: }}
+\item{\bf{BLANK: }}Blank is a count of the number of blank lines.
+
+\item{\bf{COM.L: }}Number of lines that are exclusively comments (no code).
+
+\item{\bf{COM.N: }} Number of comments in the file (a comment can be multiline).
+\end{enumerate}
+
+{\bf{\emph{Metrics Related to Complexity}}}
+\newline
+
+{\bf{McCabe's Cyclomatic Complexity}}\newline
+
 McCabe’s Cyclomatic complexity is one of the earliest complexity measures developed by Thomas J.
@@
 Every additional region in the flow graph would increase the Cyclomatic complexity by one unit.
+\begin{enumerate}
+\item{\bf{ (TOTCY)}}: Total McCabe's cyclomatic complexity (sum of all functions)
+\item{\bf{MAXCY: }} Maximum McCabe’s cyclomatic complexity. (between all functions).
+
+\item{\bf{MINCY: }} Minimum McCabe’s cyclomatic complexity.
+\item{\bf{AVGCY: }} Average McCabe’s cyclomatic complexity.
+\item{\bf{MEDCY: }} Median McCabe’s cyclomatic complexity.
+
+{\bf{Halstead's Complexity Measures}}
+
+Halstead’s complexity metrics rely on the assumption that programs should be viewed as expressions of languages both programming and written.
+It relies on the premise that there are mathematically sound relationships among the number of variables, the complexity of the code and the type of programming language statements used.
+
+\item{\bf{H.LEN: }}Halstead’s length.
+\item{\bf{H.VOL: }}Halstead's volume.
+
+\item{\bf{H.LEVEL: }}Halstead’s level.
+
+\item{\bf{H MEN.D: }}Halstead’s number of mental discriminations.
 

### 2013-12-12 paper/paper.tex
Added Data Analysis Folder with results, initial raw data plots
Added content to presentation and updated paper.tex
Added title to .tex file
Fixed title and author info
Imported result tables

@@
-\documentclass[8pt,twocolumn,a4paper]{article}
-\usepackage{amssymb, amsmath,graphicx,soul,color}
+\documentclass[9pt,twocolumn,a4paper]{article}
+\usepackage{amssymb,amsmath,graphicx,soul,color,indentfirst}
+\usepackage[affil-it]{authblk}
 
@@
 \begin{document}
+
+\author{Aldosary, Buthainah\\
+\texttt{aldosa@encs.concordia.ca}}
+\affil{Concordia University\\
+ Montreal, Canada}
+
+\date{}
+\title{Software Complexities: Do System and Change Complexity Relate?}
+
+\maketitle
+\section{Abstract}
 \section{Introduction}
 
-The complexity of a change can be very different from the complexity of the
-files or system that contain a change (\eg, a change to many simple files
-might be as complex as a change to one difficult file). Change complexity is
-related to, ``how hard is it to understand and review a change" vs system
-complexity, which measures for example, "how hard is it to understand and
-potentially modify the system".
-
-Traditional complexity measures (\eg McCabe) are normed on or require a whole
-file or system and were not designed to measure the complexity of changes and
-code fragments. The goal of this paper is to evaluate which measures are the
-best indicators of the complexity the changes to an evolving system.
+Traditional complexity measures are normalized on or require a whole file or system and were not designed to measure the complexity of code fragments.
+Therefor, the complexity of a change can be very different from the complexity of the files or system that contain a change (e.g a change to many simple files might be as complex as a change to one difficult file).
+Change complexity is related to, ``how hard is it to understand and review a change" vs system complexity, which measures for example, "how hard is it to understand and potentially modify the system".
+Traditional complexity measures do not really measure the complexity of the system but simply the size of the files that make up the system.
+As a result, they do not add any additional information beyond how large a file is.
+In contrast, all of the change complexity measures are based upon the notion that changes that are farther from each other or involve multiple entities are more complex than those closer together involving fewer entities.
+The goal of this paper is to evaluate the correlation between system complexity measures and change complexity measures of an evolving system.
+\newline
 
@@
 
+\item Do the traditional measures calculated on only the files that had changed at a certain commit correlate with the change complexity measures at the last 1000 changes?
+
 \end{enumerate}
 
-\section{Background and Measures}
+\section{Background and Motivation}
 
-\subsection{Traditional Complexity Measures}
+Traditional complexity metrics require the system as a whole and were not designed to measure the complexity of code fragments.
+McCabe’s metrics directly measures the number of linearly independent paths through a program's source code.
+Syntactic complexity metrics that are exclusively based on the structure of the program and the properties of the text (for example, redundancy of operators and operands, as Halstead’s metrics do), do not add any more information than lines of code do.
+[Isreal].
+Furthermore, multiple studies have shown that these measures correlate very strongly with lines of code and may not provide additional information about the complexity of the system.
+[54,85,61].
+As previously stated, measuring the complexity of a change differs from measuring an entire system.
+For example, a change to numerous simple functions across multiple files may be as difficult as changing the code contained within a single complex function.
+[Peter].
+In this section, we provide a background on the various measures of system and change complexity.
+
+\subsection{System (Traditional) Complexity Measures}
 {\bf{\emph{Metrics Related to Lines of Code}}}
@@
 
-\item{\bf{Source Line of Code(SLOC): }} To define SLOC we use the definition given by Conte [Conte 1986]:
-A line of code is any line of program text that is not a comment or blank line, regardless of the
-number of statements or fragments of statements on the line. This specifically includes all lines containing program headers, declarations, and executable and non-executable statements.
+\item{\bf{Source Line of Code(SLOC): }} To define SLOC, we use the definition given by Conte [Conte 1986]: A line of code is any line of program text that is not a comment of blank line, regardless of the number of statements or fragments of statements on the line.
+This specifically includes all lines containing program headers, declarations, and executable and non-executable statements.
 
-\item{\bf{Lines of Code (LOC): }} Lines of code refers to the total number of lines in each source code file including comments and blank lines.
-It is a straightforward measure and can be calculated using many available tools.
+\item{\bf{Lines of Code (LOC): }} Lines of code refers to the total number of lines in each source code file including comments and blank lines
 
@@
 Any program can be represented as a graph with the simplest element being a flat series of statements with no conditions, loops or branches.
-[1] For a graph $G$ with $n$ vertices, $e$ edges and $p$ exit points, the complexity $v$ is defined as follows:
+[Isreal] For a graph $G$ with $n$ vertices, $e$ edges and $p$ exit points, the complexity $v$ is defined as follows:
 
@@
 
-Halstead’s complexity metrics rely on the assumption that programs should be viewed as expressions of languages both programming and written.
+Halstead’s complexity metrics rely on the idea that programs should be viewed as expressions of languages both programming and written.
 It relies on the premise that there are mathematically sound relationships among the number of variables, the complexity of the code and the type of programming language statements used.
@@
 
-Each change is calcuated on the difference between two versions of a file.
-%
-All of our measures are based upon the premise that changes that are farther
-from each other or involve multiple entities are more complex than those closer
-together involving fewer entities. 
-%
-In total, we measure change complexity in seven ways:
-
-\begin{itemize}
-
-\item The churn or change size,
-
-\item The number of modified files per commit,
-
-\item The number of diffs per commit,
-
-\item The number of and distance between contiguous change blocks within a diff
-(\ie hunks),
-
-\item The directory distance between files, and
-
-\item The depth of indentation of a change.
-
-\end{itemize}
-
-%
-Since the goal of each measure is the same, to assess the complexity of a
-change, we expect some of our measures to be highly correlated. 
-%we want the simpliest most parsimonious measures. 
+Since the goal of each measure is the same, to assess the complexity of a change, we expect some of our measures to be highly correlated.
 In the interest of parsimony, we will select the simplest set of measures that adequately captures change complexity.
 
-{\bf Files and diffs:} we count the number of files contained in a commit.
+\begin{enumerate}
+
+\item{\bf{Files and diffs (Mods)}}: We count the number of files contained in a commit.
 The more files that change, the larger the affect proportion of the system.
@@
 
-{\bf Change blocks:} we measure the number of contiguous change blocks, or
-hunks, and the distance between these blocks in a modified file and sum them
-across the entire review.  Contiguous changes  are likely easier to review than
-changes that are far apart and that are potentially in multiple functions
-within a file.
+\item{\bf{Change blocks (Hunks)}}: We measure the number of contiguous change blocks, or hunks, and the distance between these blocks in a modified file and sum them across the entire review.
+Contiguous changes are likely easier to review than changes that are further apart and that are potentially in multiple functions within a file.
 
-{\bf Directory distance:} we measure the directory distance between the files
-in a commit. The premise for this measure is that files that perform similar
-functions are typically closer in the directory hierarchy than files that
-perform dissimilar functions~\cite{Bowman1999ICSE}, thus the directory
-structure loosely mirrors the system architecture~\cite{RigbySubTOSEM}. The
-distance of two files in the same directory have a distance of zero, while the
-distance for files in different directories is the number of directories
-between the two files in the directory hierarchy.  We expect that changes that
-involve files that are far apart will crosscut the functionality of a system
-and be more complex to review.
+\item{\bf{Indentation: [Hindle2008ICPC]}} creates a complexity measure that can be used to measure the complexity of the entire system or of a change by examining the level of indentation on source lines.
+We calculate their measure for reviews.
+They found that the strongest predictors of complexity were the sum and standard deviation of the indentation of source lines.
 
-{\bf Indentation:} Hindle \cite{Hindle2008ICPC} created a complexity
-measure that can be used to measure the complexity of the entire system or of a
-change by examining the level of indentation on source lines. We calculate
-their measure for reviews. They found that the strongest predictors of
-complexity were the sum and standard deviation of the indentation of source
-lines.
+\end{enumerate}
 
@@
 \section{Methodology and Data}
+
+A quantitative research methodology was used for this empirical study where the objective was to develop models pertaining to our research questions.
+The choice of this method is key as it provides a connection between the empirical observation and mathematical expressions of the quantitative relationships.
+The data used for this study was provided by Dr.
+Peter Rigby, namely from the Apache HTTP Server Project written in C.
+All measures were calculated at a particular version (commit) and the data was retrieved from June 18, 1999 – March 3, 2011; that is during a 12 year time period.
+SQL scripts were written to extract data of system and change complexity measure that answer each of the research questions using PostgreSQL 9.3.
+The resulting data was then imported into RStudio for further analysis.
+Finally, Spearman’s rank correlation coefficient matrix was deduced for each measure.
 
@@
   \centering
- %  \includegraphics[width=0.5\textwidth]{change_complexity_figure}
+  \includegraphics[width=0.5\textwidth]{change_complexity_figure}
    \caption{"Change Complexity vs. Traditional Complexity"}
@@
 \begin{equation}
-{{CCM_{v}(i)} \sim {TCM_{v(i+1)}}-{TCM_{v(i)}}}
+{{CCM_{v(i)}} = {TCM_{v(i+1)}}-{TCM_{v(i)}}}
 \end{equation}
@@
 \begin{equation}
-{{TCM_{v(i+1)}} \sum\limits_{i=0}^iCCM_{v(i)}}
+{{TCM_{v(i+1)}} = \sum\limits_{i=0}^iCCM_{v(i)}}
 \end{equation}
 
-\section{Results}
+\begin{equation}
+{{{TCM_{v(i)}} = CCM_{v(i)}}   _{n=1000}}
+\end{equation}
 
-In this section, we compare how correlated each measure is to each other.
+\section{Results and Discussion}
+
+It has been previously shown that Halstead’s length, volume, level and mental discriminations all correlate highly with each other [Israel], so it is safe to select hlen only of these metrics.
+Maxcy and sloc have also shown to correlate highly with each other so Maxcy has been chosen.
+Total cyclomatic complexity has been also added as it provides the complexity measure for the sum of all functions.
+Finally, loc has been included as it’s a measure of all text lines in the code.
+For measures pertaining to change complexity, we have chosen the simplest ones that adequately capture change complexity.
+\newline
+
+{\bf{\emph Q1) Does the difference of traditional complexity measures before and after a change correlate with change complexity measures?}}
+\newline
+
+We measure the difference between traditional complexity measures before and after a change and relate that with the change complexity measure at that particular change.
+By examining the result, we can see that measures that are related to change complexity correlate very poorly with those of system complexity (\textless 0.3).
+\newline
+
+\begin{table}[ht]
+
+\centering
+\resizebox{8cm}{!}{
+\begin{tabular}{r||rrrrrrr}
+  \hline
+ & mods & hunk\_dist & indent\_sum & loc & hlen & maxcy & totcy \\ 
+  \hline
+mods & 1.00 & 0.65 & 0.85 & 0.30 & 0.26 & 0.23 & 0.26 \\ 
+  hunk\_dist & 0.65 & 1.00 & 0.63 & 0.18 & 0.16 & 0.11 & 0.14 \\ 
+  indent\_sum & 0.85 & 0.63 & 1.00 & 0.28 & 0.24 & 0.22 & 0.24 \\ 
+  loc & 0.30 & 0.18 & 0.28 & 1.00 & 0.89 & 0.54 & 0.73 \\ 
+  hlen & 0.26 & 0.16 & 0.24 & 0.89 & 1.00 & 0.51 & 0.70 \\ 
+  maxcy & 0.23 & 0.11 & 0.22 & 0.54 & 0.51 & 1.00 & 0.64 \\ 
+  totcy & 0.26 & 0.14 & 0.24 & 0.73 & 0.70 & 0.64 & 1.00 \\ 
+   \hline
+\end{tabular}}
+\caption {\small Correlation Matrix of Research Question 1.}
+\label{tab:title} 
+\end{table}
+
+{\bf{\emph Q2)Does the weighted sum of change complexity measures up to a version correlated with the traditional complexity measures at that version? }}
+\newline
+
+We measured the last 1000 commits’ weighted sum of change complexity measures up to a single version and compared it with the traditional complexity measures of that version.
+By looking at the results in the following table, we can see that change complexity measures correlated negatively with traditional complexity measures.
+\newline
+
+\begin{table}[ht]
+\centering
+\resizebox{8cm}{!}{
+\begin{tabular}{r||rrrrrrr}
+  \hline
+ & mods & hunk\_dist & indent\_sum & loc & hlen & maxcy & totcy \\ 
+  \hline
+mods & 1.00 & 0.25 & 0.96 & -0.23 & -0.21 & -0.38 & -0.36 \\ 
+  hunk\_dist & 0.25 & 1.00 & 0.31 & -0.21 & -0.22 & -0.19 & -0.20 \\ 
+  indent\_sum & 0.96 & 0.31 & 1.00 & -0.18 & -0.16 & -0.32 & -0.30 \\ 
+  loc & -0.23 & -0.21 & -0.18 & 1.00 & 1.00 & 0.95 & 0.96 \\ 
+  hlen & -0.21 & -0.22 & -0.16 & 1.00 & 1.00 & 0.94 & 0.95 \\ 
+  maxcy & -0.38 & -0.19 & -0.32 & 0.95 & 0.94 & 1.00 & 1.00 \\ 
+  totcy & -0.36 & -0.20 & -0.30 & 0.96 & 0.95 & 1.00 & 1.00 \\ 
+   \hline
+\end{tabular}}
+\caption {\small Correlation Matrix of Research Question 2.}
+\label{tab:title} 
+\end{table}
+
+{\bf{\emph Q3)Do the traditional measures calculated on only the files that had changed at a certain commit correlate with the change complexity measures at the last 1000 changes?}}
+\newline
+
+To answer the third research question, first we identified the files that have undergone changes in a version and extracted the traditional measures for them.
+Then, we measured the change complexity measures for the last 1000 changes and compared these two measures.
+Table 3 (???) shows the correlation matrix which shows that there is a very poor correlation between the change complexity measures and traditional complexity measures (\textless 0.3).
+
+\begin{table}[ht]
+
+\centering
+\resizebox{8cm}{!}{
+\begin{tabular}{r||rrrrrrr}
+  \hline
+ & mods & hunk\_dist & indent\_sum & loc & hlen & maxcy & totcy \\ 
+  \hline
+mods & 1.00 & 0.66 & 0.86 & 0.23 & 0.23 & 0.21 & 0.23 \\ 
+  hunk\_dist & 0.66 & 1.00 & 0.63 & 0.27 & 0.26 & 0.22 & 0.28 \\ 
+  indent\_sum & 0.86 & 0.63 & 1.00 & 0.26 & 0.26 & 0.28 & 0.27 \\ 
+  loc & 0.23 & 0.27 & 0.26 & 1.00 & 0.99 & 0.82 & 0.96 \\ 
+  hlen & 0.23 & 0.26 & 0.26 & 0.99 & 1.00 & 0.85 & 0.96 \\ 
+  maxcy & 0.21 & 0.22 & 0.28 & 0.82 & 0.85 & 1.00 & 0.84 \\ 
+  totcy & 0.23 & 0.28 & 0.27 & 0.96 & 0.96 & 0.84 & 1.00 \\ 
+   \hline
+\end{tabular}}
+\caption {\small Correlation Matrix of Research Question 3.}
+\label{tab:title} 
+\end{table}
 

### 2013-12-13 paper/paper.tex
Final latex paper to be submitted

@@
 \usepackage[affil-it]{authblk}
+\usepackage[T1]{fontenc}
 
@@
 \date{}
-\title{Software Complexities: Do System and Change Complexity Relate?}
+\title{\bf{Software Complexities: Do System and Change Complexity Relate?}}
 
@@
 \section{Abstract}
+{\em System complexity is a term that encompasses the internal interations of a software and describes them.
+As the number of these entities increase, the number of interactions would exponentially increase leading to an unmanageable complexity.
+On the other hand, change complexity encompasses the artifacts that make up a single change such as diffs, patches and revisions.
+To see how well these two measures relate, we extracted data from the Apache HTTP Server Open Source Project and calculated the correlations between pairs of metrics.
+Upon completion of the analysis, we found that system and change complexity metrics correlate poorly.
+This is due to the fact that they exclusively measure different things.
+
+}
 \section{Introduction}
@@
 
+The rest of the paper is organized as follows: Section 3 describes the background and motivation of this work.
+Section 4 describes the method and data used.
+Section 5 and 6 present the results and discussion as well as threats to validity.
+Section 7 is the conclusion and finally in section 8, we describe potential future work.
+
 \section{Background and Motivation}
@@
 Traditional complexity metrics require the system as a whole and were not designed to measure the complexity of code fragments.
-McCabe’s metrics directly measures the number of linearly independent paths through a program's source code.
-Syntactic complexity metrics that are exclusively based on the structure of the program and the properties of the text (for example, redundancy of operators and operands, as Halstead’s metrics do), do not add any more information than lines of code do.
-[Isreal].
+McCabe's metrics directly measures the number of linearly independent paths through a program's source code.
+Syntactic complexity metrics that are exclusively based on the structure of the program and the properties of the text (for example, redundancy of operators and operands, as Halstead's metrics do), do not add any more information than lines of code do \cite {Israel}.
 Furthermore, multiple studies have shown that these measures correlate very strongly with lines of code and may not provide additional information about the complexity of the system.
-[54,85,61].
+\cite {Graves, Leszak, Herraiz}.
 As previously stated, measuring the complexity of a change differs from measuring an entire system.
-For example, a change to numerous simple functions across multiple files may be as difficult as changing the code contained within a single complex function.
-[Peter].
+For example, a change to numerous simple functions across multiple files may be as difficult as changing the code contained within a single complex function \cite{Peter}.
 In this section, we provide a background on the various measures of system and change complexity.
@@
 
-\item{\bf{BLANK: }}Blank is a count of the number of blank lines.
+\item{\bf{\textsc{Blank: }}}Blank is a count of the number of blank lines.
 
-\item{\bf{COM.L: }}Number of lines that are exclusively comments (no code).
+\item{\bf{\textsc{Com.l: }}}Number of lines that are exclusively comments (no code).
 
-\item{\bf{COM.N: }} Number of comments in the file (a comment can be multiline).
+\item{\bf{\textsc{Com.n: }}} Number of comments in the file (a comment can be multiline).
 \end{enumerate}
@@
 
-McCabe’s Cyclomatic complexity is one of the earliest complexity measures developed by Thomas J.
+McCabe's Cyclomatic complexity is one of the earliest complexity measures developed by Thomas J.
 McCabe in 1976.
 It directly measures the number of linearly independent paths through a program's source code.
-Any program can be represented as a graph with the simplest element being a flat series of statements with no conditions, loops or branches.
-[Isreal] For a graph $G$ with $n$ vertices, $e$ edges and $p$ exit points, the complexity $v$ is defined as follows:
+Any program can be represented as a graph with the simplest element being a flat series of statements with no conditions, loops or branches \cite{Israel}.
+For a graph $G$ with $n$ vertices, $e$ edges and $p$ exit points, the complexity $v$ is defined as follows:
 
@@
 \begin{enumerate}
-\item{\bf{ (TOTCY)}}: Total McCabe's cyclomatic complexity (sum of all functions)
-\item{\bf{MAXCY: }} Maximum McCabe’s cyclomatic complexity. (between all functions).
+\item{\bf{\textsc{Totcy}}} Total McCabe's cyclomatic complexity (sum of all functions)
+\item{\bf{\textsc{Maxcy}}} Maximum McCabe's cyclomatic complexity. (between all functions).
 
-\item{\bf{MINCY: }} Minimum McCabe’s cyclomatic complexity.
-\item{\bf{AVGCY: }} Average McCabe’s cyclomatic complexity.
-\item{\bf{MEDCY: }} Median McCabe’s cyclomatic complexity.
+\item{\bf{\textsc{Mincy}}} Minimum McCabe's cyclomatic complexity.
+\item{\bf{\textsc{Avgcy}}} Average McCabe's cyclomatic complexity.
+\item{\bf{\textsc{Medcy}}} Median McCabe's cyclomatic complexity.
 
@@
 
-Halstead’s complexity metrics rely on the idea that programs should be viewed as expressions of languages both programming and written.
+Halstead's complexity metrics rely on the idea that programs should be viewed as expressions of languages both programming and written.
 It relies on the premise that there are mathematically sound relationships among the number of variables, the complexity of the code and the type of programming language statements used.
 
-\item{\bf{H.LEN: }}Halstead’s length.
-\item{\bf{H.VOL: }}Halstead's volume.
+\item{\bf{\textsc{Hlen}}} Halstead's length.
+\item{\bf{\textsc{Hvol}}} Halstead's volume.
 
-\item{\bf{H.LEVEL: }}Halstead’s level.
+\item{\bf{\textsc{Hlevel}}} Halstead's level.
 
-\item{\bf{H MEN.D: }}Halstead’s number of mental discriminations.
+\item{\bf{\textsc{H.Men.D}}} Halstead's number of mental discriminations.
 
@@
 
-\item{\bf{Indentation: [Hindle2008ICPC]}} creates a complexity measure that can be used to measure the complexity of the entire system or of a change by examining the level of indentation on source lines.
+\item{\bf{Indentation}}: Hindle {\em {et al.}} \cite{Hindle} created a complexity measure that can be used to measure the complexity of the entire system or of a change by examining the level of indentation on source lines.
 We calculate their measure for reviews.
@@
 Peter Rigby, namely from the Apache HTTP Server Project written in C.
-All measures were calculated at a particular version (commit) and the data was retrieved from June 18, 1999 – March 3, 2011; that is during a 12 year time period.
+All measures were calculated at a particular version (commit) and the data was retrieved from June 18, 1999 - March 3, 2011; that is during a 12 year time period.However, no further information is given if this data includes header files or not.
 SQL scripts were written to extract data of system and change complexity measure that answer each of the research questions using PostgreSQL 9.3.
 The resulting data was then imported into RStudio for further analysis.
-Finally, Spearman’s rank correlation coefficient matrix was deduced for each measure.
+Finally, Spearman's rank correlation coefficient matrix was deduced for each measure.
 
-In figure 1 we illustrate a hypothetical system which undergoes numerous changes.
+In figure \ref{fig:ccVstc} we illustrate a system which undergoes numerous changes.
 The vertical lines represent each version.
@@
    \caption{"Change Complexity vs. Traditional Complexity"}
+   \label{fig:ccVstc}
 \end{figure}
@@
 
-It has been previously shown that Halstead’s length, volume, level and mental discriminations all correlate highly with each other [Israel], so it is safe to select hlen only of these metrics.
+It has been previously shown that Halstead's length, volume, level and mental discriminations all correlate highly with each other \cite{Israel}, so it is safe to select hlen only of these metrics.
 Maxcy and sloc have also shown to correlate highly with each other so Maxcy has been chosen.
 Total cyclomatic complexity has been also added as it provides the complexity measure for the sum of all functions.
-Finally, loc has been included as it’s a measure of all text lines in the code.
+Finally, loc has been included as it is a measure of all text lines in the code.
 For measures pertaining to change complexity, we have chosen the simplest ones that adequately capture change complexity.
@@
 \caption {\small Correlation Matrix of Research Question 1.}
-\label{tab:title} 
+\label{tab:table1}
+
 \end{table}
@@
 
-We measured the last 1000 commits’ weighted sum of change complexity measures up to a single version and compared it with the traditional complexity measures of that version.
+We measured the last 1000 commits' weighted sum of change complexity measures up to a single version and compared it with the traditional complexity measures of that version.
 By looking at the results in the following table, we can see that change complexity measures correlated negatively with traditional complexity measures.
@@
 \caption {\small Correlation Matrix of Research Question 2.}
-\label{tab:title} 
+\label{tab:table2} 
 \end{table}
@@
 Then, we measured the change complexity measures for the last 1000 changes and compared these two measures.
-Table 3 (???) shows the correlation matrix which shows that there is a very poor correlation between the change complexity measures and traditional complexity measures (\textless 0.3).
+Table \ref{tab:table3} shows the correlation matrix which shows that there is a very poor correlation between the change complexity measures and traditional complexity measures (\textless 0.3).
 
@@
 \caption {\small Correlation Matrix of Research Question 3.}
-\label{tab:title} 
+
+\label{tab:table3} 
 \end{table}
 
+From the previous results we have identified that traditional and change complexity measures correlate very poorly.
+This is due to the nature of how these measurements are extracted and what they concretely measure.
+For example, change complexity measures are extracted from diffs, revisions and patches as opposed to traditional complexity measures, which are extracted from the source code as a whole \cite{Hindle}.
+We can therefor explain the poor correlation with the fact that these two complexity metrics measure different things
+
 \section{Threats to validity}
-
+In our study we calculated the complexity measures on each particular commits (aka.
+each version), however the correlation between traditional complexity measures in Israel's study \cite{Israel} was done at a file level.
+Israel also made account of the confounding effect of file size.
+The result was that even though file size influences the values of the correlations, it does not substantially change the conclusion extracted from those correlations in any case.
+Therefor, in order to rule out this effect, the total size of the system must be taken into consideration.
+Another threat to validity is the source of inconsistencies in measuring traditional complexity metrics between header files and non-header files.
 \section{Conclusion}
 
-Future work: Does McCabe predict anything? Do the change complexity measures do a better job of prediction?
+In this work we have chosen the Apache HTTP Server Open Source Project and extracted data related to traditional and change complexity measures within a 12-year period.
+We have shown that these two measures correlate poorly because they measure different things in different contexts.
 
+\section{Future Work}
+Future research directions include using traditional and change complexity metrics as an indication of the maintainability and evolvability of a system; such as productivity and number of bugs.
+
+\begin{thebibliography}{9}
+
+\bibitem{Israel}
+I. Herraiz, A. Hassan  {\em Beyond Lines of Code: Do We Need More Complexity Metrics?} Making Software. Greg Wilson and Andy Oram,
+  O'Reilly Media, Inc., 
+  2010.
+
+\bibitem{Peter}
+P. Rigby (2011), {\em Understanding Open Source Software Peer Review:
+Review Processes, Parameters and Statistical Models, and
+Underlying Behaviours and Mechanisms.},  
+Ph.D. Thesis. University of Victoria, Canada.
+
+\bibitem{Hindle}
+A. Hindle, M. W. Godfrey, R. C. Holt. {\em Reading Beside the Lines: Indentation as a Proxy for Complexity Metrics.} In ICPC: Proceedings of the 16th IEEE International
+Conference on Program Comprehension, pages 133-142. IEEE Computer Society,
+2008.
+
+\bibitem{Graves}
+T. L. Graves, A. F. Karr, J. S. Marron, and H. Siy. {\em Predicting fault incidence using
+software change history.} IEEE Trans. Softw. Eng., 26(7):653-661, 2000.
+
+\bibitem{Leszak}
+M. Leszak, D. E. Perry, and D. Stoll. { \em Classification and evaluation of defects in a
+project retrospective.} Journal of Systems and Software, 61(3):173-187, 2002.
+
+\bibitem{Herraiz}
+I. Herraiz, J. M. Gonzalez-Barahona, and G. Robles. { \em Towards a theoretical model
+for software growth.} In MSR: Proceedings of the Fourth International Workshop on
+Mining Software Repositories, pages 21-28. IEEE Computer Society, 2007.
+
+\end{thebibliography}
 \end{document}

### 2013-12-18 paper/paper.tex
Added timeseries plots

@@
 \author{Aldosary, Buthainah\\
-\texttt{aldosa@encs.concordia.ca}}
+\texttt{b_aldosa@encs.concordia.ca}}
 \affil{Concordia University\\
@@
 Peter Rigby, namely from the Apache HTTP Server Project written in C.
-All measures were calculated at a particular version (commit) and the data was retrieved from June 18, 1999 - March 3, 2011; that is during a 12 year time period.However, no further information is given if this data includes header files or not.
+All measures were calculated at a particular version (commit) and the data was retrieved from June 18, 1999 - March 3, 2011; that is during a 12 year time period.
+However, no further information is given if this data includes header files or not.
 SQL scripts were written to extract data of system and change complexity measure that answer each of the research questions using PostgreSQL 9.3.
 The resulting data was then imported into RStudio for further analysis.
-Finally, Spearman's rank correlation coefficient matrix was deduced for each measure.
+Finally, Spearman's rank correlation coefficient matrix was deduced to answer each of our research questions.
 
@@
 
+%Traditional Complexity timeseries
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{TradComplexityMeasuresPlot}
+   \caption{Traditional Complexity Measures Timeseries }
+   \label{fig:tct}
+\end{figure}
+
+%Change Complexity timeseries
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{ChangeComplexMeasuresPlot}
+   \caption{Change Complexity Measures Timeseries }
+   \label{fig:cct}
+\end{figure}
+
+%Change and Traditional Complexity timeseries
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{ChangeandTradPlot}
+   \caption{Traditional and Change Complexity Measures Timeseries }
+   \label{fig:TandCCT}
+\end{figure}
+
 \section{Threats to validity}

### 2013-12-18 paper/paper.tex
.

@@
 \author{Aldosary, Buthainah\\
-\texttt{b_aldosa@encs.concordia.ca}}
+\texttt{aldosa@encs.concordia.ca}}
 \affil{Concordia University\\
@@
 These are denoted by CCM and TCM respectively.
+To answer our research questions, for example at version 3, we want to see if \textsc{ ccm(v3)= tcm(v4)-tcm(v3)} (\ref{eq:Eq2}), \textsc {tcm(v4)=ccm(v1)+ccm(v2)+ccm(v3)
+} (\ref{eq:Eq3}), finally we want to see if \textsc {tcm(v3)=ccm(v3)} for the last 1000 commits (\ref{eq:Eq4}).
 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:Eq2}
 {{CCM_{v(i)}} = {TCM_{v(i+1)}}-{TCM_{v(i)}}}
@@
 
-\begin{equation}
+\begin{equation} \label{eq:Eq3}
 {{TCM_{v(i+1)}} = \sum\limits_{i=0}^iCCM_{v(i)}}
@@
 
-\begin{equation}
+\begin{equation}  \label{eq:Eq4}
 {{{TCM_{v(i)}} = CCM_{v(i)}}   _{n=1000}}

### 2013-12-19 paper/paper.tex
Final paper

@@
 \author{Aldosary, Buthainah\\
-\texttt{aldosa@encs.concordia.ca}}
+\texttt{b\_aldosa@encs.concordia.ca}}
 \affil{Concordia University\\
@@
 
-%TODO weighted number of people who changed a file?
-
 \section{Methodology and Data}
@@
 We measured the last 1000 commits' weighted sum of change complexity measures up to a single version and compared it with the traditional complexity measures of that version.
-By looking at the results in the following table, we can see that change complexity measures correlated negatively with traditional complexity measures.
+By looking at the results in table \ref{tab:table2}, we can see that change complexity measures correlated negatively with traditional complexity measures.
 \newline
@@
 \caption {\small Correlation Matrix of Research Question 3.}
-
 \label{tab:table3} 
@@
 
-From the previous results we have identified that traditional and change complexity measures correlate very poorly.
+From the previous results of each of our research questions, we can see that syntactic complexity metrics don't provide more complexity information than lines of code do.
+We have also identified that traditional and change complexity measures correlate very poorly.
 This is due to the nature of how these measurements are extracted and what they concretely measure.
 For example, change complexity measures are extracted from diffs, revisions and patches as opposed to traditional complexity measures, which are extracted from the source code as a whole \cite{Hindle}.
-We can therefor explain the poor correlation with the fact that these two complexity metrics measure different things
+We can therefor explain the poor correlation with the underlying artifacts that these metrics are based on.
+Therefor, these complexity measures can not be substituted to be used as a proxy for the amount of effort that is needed to comprehend a piece of code.
+\newline
+A snapshot of the project from it's start date until it's end date is shown in the next figures.
+The traditional complexity measures is shown in Figure \ref{fig:tct} in isolation, namely Locs, Totcy and Hlen.
+We can see that they all follow an increasing trend with Hlen and Loc decreasing at the same points in time.\newline
+The change complexity measures (Mods, Hunk dist and Indent Sum) is shown in Figure \ref{fig:cct} also in isolation.
+They too follow a same trend with respect to each other.
+Finally, figure \ref{fig:TandCCT} shows the totCy and Hunk dist.
+In this case, we can see that they follow very different trends.
+This confirms that these two metrics indeed measure different things.
 
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{TradComplexityMeasuresPlot}
+  \includegraphics[width=0.5\textwidth]
+  {TradPlot}
    \caption{Traditional Complexity Measures Timeseries }
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{ChangeComplexMeasuresPlot}
+  \includegraphics[width=0.5\textwidth]
+  {ChangePlot}
    \caption{Change Complexity Measures Timeseries }
@@
 
-\section{Threats to validity}
+\section{Threats to Validity}
 In our study we calculated the complexity measures on each particular commits (aka.
@@
 In this work we have chosen the Apache HTTP Server Open Source Project and extracted data related to traditional and change complexity measures within a 12-year period.
-We have shown that these two measures correlate poorly because they measure different things in different contexts.
+For each of our research questions, we have shown that
+traditional complexity measures, in particular those that are exclusively based on the structure of the program or properties of text, do not provide additional information on the amount of effort that's required to fully understand a piece of code.
+Therefor, lines of code are sufficient to comprehend code fragments.
+We have also shown that, for each research question, while change complexity measures correlate very strongly with one another, the same result is not the case with traditional complexity measures (in some cases, there is a negative relationship).
+This is because they measure different things in different contexts and therefor results in a weak relationship between them.
 
